---
title: "CIN signatures linked to changes in expression of genes involved in DNA repair"
output:
  html_document:
    df_print: paged
---

#### LOAD PACKAGES & ENVIRONMENT
```{r, warning=FALSE}
options(java.parameters = "-Xmx8000m")

library(dplyr)
library(reshape2)
library(ggplot2)
library(ppcor)
library(ggrepel)
library(ggpubr)
library(fgsea)
library(ggplot2)
library(ggthemes)
library(lemon)
library(RColorBrewer)
library(patchwork)

#PATHS
BASE=dirname(rstudioapi::getSourceEditorContext()$path)
OUT=file.path(BASE, "output")
```

##SELECT GENE EXPRESSION DATA OF DNA REPAIR GENES
```{r}
# #Performed in the cluster
# #List of DNA repair genes downloaded from Reactome (DNA repair pathway) --> included in the CNSignatures/0_external_data folder

# #Directory
# INPUT_DIR <- "/storage/scratch01/users/bhernando"

# #Read DNArepair genes from reactome txt
# DNARepair<-read.delim(paste0(INPUT_DIR,"/DNARepair_genes.txt"), head=T)

# #Load TCGA gene expression
# TCGA_expression<-read.delim(paste0(INPUT_DIR,"/CosmicCompleteGeneExpression.tsv.gz"), head=T) --> Download from https://data.broadinstitute.org/snowman/ofer/COSMIC/

# #Select expression of DNA repair genes
# DNArepair_expression<-TCGA_expression[TCGA_expression[,3] %in% DNARepair$gene, ]

# #Output
# write.table(DNArepair_expression, file="DNARepair_expression.txt", sep="\t", row.names=F)
```


##LOAD DATA 
```{r}
##Load data linking genes and DNA repair mechanisms
DNARepairMechanisms <- read.delim(paste0(BASE,"/input/DNARepairMechanisms.txt"), sep="\t", head=T)

##Load TCGA CIN signature activities
TCGA_signatures<-readRDS(paste0(BASE,"/input/Signature_Compendium_v5_Cosine-0.74_Activities_THRESH95_NAMESAPRIL21.rds"))

##Load TCGA expression of genes involved in DNA repair
# This file is too large for github --> please contact us if needed. The input matrix only includes expression data of DNA Repair genes from COSMIC. 
# The matrix includes z-scores of all samples (rows) for all genes (columns)
TCGA_repairexpression <- read.delim(paste0(BASE,"/input/DNARepair_expression.txt"), head=T) 
TCGA_repairexpression <- acast(TCGA_repairexpression, SAMPLE_NAME~GENE_NAME, value.var="Z_SCORE")

#Match samples with CIN signatures & gene expression data
#Use 12 digit code of TCGA samples
rownames(TCGA_signatures)<-substr(row.names(TCGA_signatures),1,12)
rownames(TCGA_repairexpression) <- substr(row.names(TCGA_repairexpression),1,12)
TCGA_repairexpression_sign <- TCGA_repairexpression[rownames(TCGA_repairexpression) %in% rownames(TCGA_signatures),]
TCGA_signatures_exp <- TCGA_signatures[match(rownames(TCGA_repairexpression_sign),rownames(TCGA_signatures)),]

#Summary
print("Total number of TCGA samples with signatures and gene expression data:")
nrow(TCGA_signatures_exp)

print("Total number of DNA repair genes:")
ncol(TCGA_repairexpression_sign)
```

##CORRELATION ANALYSES BETWEEN SIGNATURES AND EXPRESSION OF DNA REPAIR GENES
```{r, warning=FALSE}
#Pearson correlations between CIN activities and gene expression (z.scores)
correlations<-c()
for(i in 1:ncol(TCGA_signatures_exp))
{
  print(paste0("Correlations for signature ",colnames(TCGA_signatures_exp)[i]))
  sig_cor<-c()
  for(j in 1:ncol(TCGA_repairexpression_sign))
  {
    cor<-cor.test(TCGA_repairexpression_sign[,j],TCGA_signatures_exp[,i],method="spearman",alternative="two.sided")
    sig_cor<-rbind(sig_cor,c(colnames(TCGA_signatures_exp)[i],colnames(TCGA_repairexpression_sign)[j],
                               cor$estimate,cor$p.value))
  }
  sig_cor<-data.frame(sig_cor,stringsAsFactors = F)
  sig_cor<-cbind(sig_cor,q.value=p.adjust(sig_cor[,4],method="BH"))
  correlations<-rbind(correlations,sig_cor)
}

colnames(correlations) <- c("signature","gene","rho","p.value","q.value")
correlations <- data.frame(correlations,stringsAsFactors = F)
correlations <- correlations[order(correlations$signature,correlations$q.value),]
correlations_filt<-correlations[correlations$q.value<0.05,]

#Save
write.table(correlations, paste0(OUT,"/Correlations_SignExpression.txt"),sep="\t", row.names=F)

#Summary
print("Number of significant correlations:")
table(correlations_filt$signature)
```


##CORRELATION ANALYSES BETWEEN IHR AND NON-IHR SIGNATURES AND EXPRESSION OF DNA REPAIR GENES
```{r, warning=FALSE}
#Select IHR signatures
signatures <- as.data.frame(TCGA_signatures_exp[,c(2,3,5)])

#Extract Non-IHR exposures
signatures$No.IHR <- 1-rowSums(signatures)

#Spearson correlations
correlations<-c()
for(i in 1:ncol(signatures))
{
  print(paste0("Correlations for signature ",colnames(signatures)[i]))
  sig_cor<-c()
  for(j in 1:ncol(TCGA_repairexpression_sign))
  {
    cor<-cor.test(TCGA_repairexpression_sign[,j],signatures[,i],method="spearman",alternative = "two.sided")
    sig_cor<-rbind(sig_cor,c(colnames(signatures)[i],colnames(TCGA_repairexpression_sign)[j],
                               cor$estimate,cor$p.value))
  }
  sig_cor<-data.frame(sig_cor,stringsAsFactors = F)
  sig_cor<-cbind(sig_cor,q.value=p.adjust(sig_cor[,4],method="BH"))
  correlations<-rbind(correlations,sig_cor)
}

colnames(correlations) <- c("signature","gene","rho","p.value","q.value")
correlations <- data.frame(correlations,stringsAsFactors = F)
correlations <- correlations[order(correlations$signature,correlations$q.value),]
correlations_filt<-correlations[correlations$q.value<0.05,]

#Save
write.table(paste0(OUT,"/Correlations_SignIHRExpression.txt"),sep="\t", row.names=F)

#Summary
print("Number of significant correlations:")
table(correlations_filt$signature)
```

###GENES DIFFERENTIALY EXPRESSED IHR VS NON-IHR
```{r}
#Sign data scaled
mPCAWGScaled = scale(TCGA_signatures_exp)
dfPS <-data.frame(mPCAWGScaled[,c("CX2", "CX3", "CX5")])
dfPS$Class = "NO.IHR"
dfPS$Class[ dfPS$CX2 > 1.25 & dfPS$CX2 > dfPS$CX3 & dfPS$CX2 > dfPS$CX5 ] = "IHR"
dfPS$Class[ dfPS$CX3 > 1.25 & dfPS$CX3 > dfPS$CX2 & dfPS$CX3 > dfPS$CX5 ] = "IHR"
dfPS$Class[ dfPS$CX5 > 1.25 & dfPS$CX5 > dfPS$CX2 & dfPS$CX5 > dfPS$CX3 ] = "IHR"
dfPS$SAMPLE_NAME <- rownames(TCGA_signatures_exp)

#Link genes with DNA repair pathway --> too large to be included
geneset_expression <- read.delim(paste0(BASE,"/input/DNARepair_expression.txt"), head=T)
geneset_expression$SAMPLE_NAME <- substr(geneset_expression$SAMPLE_NAME,1,12)
geneset_expression <- geneset_expression[geneset_expression$SAMPLE_NAME %in% dfPS$SAMPLE_NAME, ]

#Join gene expression and signatures activities
geneset_expression <- left_join(geneset_expression, dfPS, by="SAMPLE_NAME")
geneset_expression <- geneset_expression[!is.na(geneset_expression$Class),]

genes <- unique(geneset_expression$GENE_NAME)
i=1
diffgenes <- c()
for (i in 1:length(genes)){
  gene <- genes[i]
  dat <- geneset_expression[geneset_expression$GENE_NAME %in% gene, ]
  ttest <- t.test(Z_SCORE~Class, dat, alternative="two.sided")
  res <-cbind(gene,ttest$p.value)
  diffgenes <- rbind(diffgenes, res)
}
diffgenes<-data.frame(diffgenes,stringsAsFactors = F)
colnames(diffgenes)[2]<-c("pvalue")
diffgenes$pvalue <- as.numeric(diffgenes$pvalue)

#Gene expressed significantly different between IHR and Non-IHR 
diffgenes_IHR <- diffgenes[diffgenes$pvalue<0.01, ]

```

###SUPPLEMENTARY VOLCANOPLOTS --> Supplementary Figure 12
#Non-NER pathways
```{r}
correlations <- read.delim(paste0(OUT,"/Correlations_SignIHRExpression.txt"), sep="\t", head=T)
correlations$rho <- as.numeric(correlations$rho)

#select only genes differentialy expressed between IHR and Non-IHR samples
correlations <- correlations[correlations$gene %in% diffgenes_IHR$gene,]

#Volcano plots per pathway
xmax=max(correlations$rho)
xmin=min(correlations$rho)
ymax=max(-log10(correlations$q.value))

#alt-NHEJ
path   <- "alt-NHEJ"
genes  <- DNARepairMechanisms$gene[DNARepairMechanisms$DNARepairMechanism %in% path]
cor    <- correlations[correlations$gene %in% genes,]
cor    <- mutate(cor, significant=ifelse(abs(cor$rho)<0.1, "No significant", 
                                           ifelse(abs(cor$rho)>0.1 & cor$rho>0, "Positive",
                                                  "Negative")))
p1 <- ggplot(cor[cor$signature=="CX2",], aes(x=as.numeric(rho), y=-log10(q.value), label = gene)) +      
    geom_point(aes(fill=significant, color=significant), show.legend=F, size=1)+
    geom_text_repel(aes(label = ifelse(-log10(q.value) > 25, gene, "")), size=1.8, max.overlaps = Inf,min.segment.length = 0)+
    theme_classic () +
    scale_color_manual(values=c("Positive"="palegreen1","Negative"="red1","No significant"="grey50"),
                   labels = c("Positive","Negative","No significant"))+
    labs(title=path, x="rho", y = "-log10(q.value)")+
    geom_text(x=0, y=ymax-1, label="CX2", size=3, colour="black")+
    ylim(0,ymax)+
    xlim(xmin, xmax)+
    coord_capped_cart(left = "both", bottom = "both") +
    theme(axis.title.y = element_text(margin = margin(r=0.1, unit = "cm"),size=6))+
    theme(axis.title.x = element_text(margin = margin(t=0.1, unit = "cm"),size=6))+
    theme(axis.text.x = element_text(colour = "black", size = 6))+
    theme(axis.text.y = element_text(colour = "black", size = 6))+
    theme(title = element_text(face="bold", colour = "black", size = 7))+
    theme(plot.margin = margin(t = 0.2, r = 0.2, b = 0.2, l = 0.2, unit = "cm"))

p2 <- ggplot(cor[cor$signature=="CX5",], aes(x=as.numeric(rho), y=-log10(q.value), label = gene)) +      
    geom_point(aes(fill=significant, color=significant), show.legend=F, size=1)+
    geom_text_repel(aes(label = ifelse(-log10(q.value) > 25, gene, "")), size=1.8, max.overlaps = Inf,min.segment.length = 0)+
    theme_classic () +
    scale_color_manual(values=c("Positive"="palegreen1","Negative"="red1","No significant"="grey50"),
                   labels = c("Positive","Negative","No significant"))+
    labs(title="", x="rho", y = "-log10(q.value)")+
    geom_text(x=0, y=ymax-1, label="CX5", size=3, colour="black")+
    ylim(0,ymax)+
    xlim(xmin, xmax)+
    coord_capped_cart(left = "both", bottom = "both") +
    theme(axis.title.y = element_text(margin = margin(r=0.1, unit = "cm"),size=6))+
    theme(axis.title.x = element_text(margin = margin(t=0.1, unit = "cm"),size=6))+
    theme(axis.text.x = element_text(colour = "black", size = 6))+
    theme(axis.text.y = element_text(colour = "black", size = 6))+
    theme(title = element_text(face="bold", colour = "black", size = 7))+
    theme(plot.margin = margin(t = 0.2, r = 0.2, b = 0.2, l = 0.2, unit = "cm"))

p3 <- ggplot(cor[cor$signature=="CX3",], aes(x=as.numeric(rho), y=-log10(q.value), label = gene)) +
    geom_point(aes(fill=significant, color=significant), show.legend=F, size=1)+
    geom_text_repel(aes(label = ifelse(-log10(q.value) > 25, gene, "")), size=1.8, max.overlaps = Inf,min.segment.length = 0)+
    theme_classic () +
    scale_color_manual(values=c("Positive"="palegreen1","Negative"="red1","No significant"="grey50"),
                   labels = c("Positive","Negative","No significant"))+
    labs(title="", x="rho", y = "-log10(q.value)")+
    geom_text(x=0, y=ymax-1, label="CX3", size=3, colour="black")+
    ylim(0,ymax)+
    xlim(xmin, xmax)+
    coord_capped_cart(left = "both", bottom = "both") +
    theme(axis.title.y = element_text(margin = margin(r=0.1, unit = "cm"),size=6))+
    theme(axis.title.x = element_text(margin = margin(t=0.1, unit = "cm"),size=6))+
    theme(axis.text.x = element_text(colour = "black", size = 6))+
    theme(axis.text.y = element_text(colour = "black", size = 6))+
    theme(title = element_text(face="bold", colour = "black", size = 7))+
    theme(plot.margin = margin(t = 0.2, r = 0.2, b = 0.2, l = 0.2, unit = "cm"))
  
pp <- ggarrange(p1, p2, p3, align = "h", ncol = 3, nrow = 1)

#BER 
path   <- "BER"
genes  <- DNARepairMechanisms$gene[DNARepairMechanisms$DNARepairMechanism %in% path]
cor    <- correlations[correlations$gene %in% genes,]
cor    <- mutate(cor, significant=ifelse(abs(cor$rho)<0.1, "No significant", 
                                           ifelse(abs(cor$rho)>0.1 & cor$rho>0, "Positive",
                                                  "Negative")))
p1 <- ggplot(cor[cor$signature=="CX2",], aes(x=as.numeric(rho), y=-log10(q.value), label = gene)) +      
    geom_point(aes(fill=significant, color=significant), show.legend=F, size=1)+
    geom_text_repel(aes(label = ifelse(-log10(q.value) > 25, gene, "")), size=1.8, max.overlaps = Inf,min.segment.length = 0)+
    theme_classic () +
    scale_color_manual(values=c("Positive"="palegreen1","Negative"="red1","No significant"="grey50"),
                   labels = c("Positive","Negative","No significant"))+
    labs(title=path, x="rho", y = "-log10(q.value)")+
    geom_text(x=0, y=ymax-5, label="CX2", size=3, colour="black")+
    ylim(0,ymax)+
    xlim(xmin, xmax)+
    coord_capped_cart(left = "both", bottom = "both") +
    theme(axis.title.y = element_text(margin = margin(r=0.1, unit = "cm"),size=6))+
    theme(axis.title.x = element_text(margin = margin(t=0.1, unit = "cm"),size=6))+
    theme(axis.text.x = element_text(colour = "black", size = 6))+
    theme(axis.text.y = element_text(colour = "black", size = 6))+
    theme(title = element_text(face="bold", colour = "black", size = 7))+
    theme(plot.margin = margin(t = 0.2, r = 0.2, b = 0.2, l = 0.2, unit = "cm"))

p2 <- ggplot(cor[cor$signature=="CX5",], aes(x=as.numeric(rho), y=-log10(q.value), label = gene)) +      
    geom_point(aes(fill=significant, color=significant), show.legend=F, size=1)+
    geom_text_repel(aes(label = ifelse(-log10(q.value) > 25, gene, "")), size=1.8, max.overlaps = Inf,min.segment.length = 0)+
    theme_classic () +
    scale_color_manual(values=c("Positive"="palegreen1","Negative"="red1","No significant"="grey50"),
                   labels = c("Positive","Negative","No significant"))+
    labs(title="", x="rho", y = "-log10(q.value)")+
    geom_text(x=0, y=ymax-5, label="CX5", size=3, colour="black")+
    ylim(0,ymax)+
    xlim(xmin, xmax)+
    coord_capped_cart(left = "both", bottom = "both") +
    theme(axis.title.y = element_text(margin = margin(r=0.1, unit = "cm"),size=6))+
    theme(axis.title.x = element_text(margin = margin(t=0.1, unit = "cm"),size=6))+
    theme(axis.text.x = element_text(colour = "black", size = 6))+
    theme(axis.text.y = element_text(colour = "black", size = 6))+
    theme(title = element_text(face="bold", colour = "black", size = 7))+
    theme(plot.margin = margin(t = 0.2, r = 0.2, b = 0.2, l = 0.2, unit = "cm"))

p3 <- ggplot(cor[cor$signature=="CX3",], aes(x=as.numeric(rho), y=-log10(q.value), label = gene)) +
    geom_point(aes(fill=significant, color=significant), show.legend=F, size=1)+
    geom_text_repel(aes(label = ifelse(-log10(q.value) > 25, gene, "")), size=1.8, max.overlaps = Inf,min.segment.length = 0)+
    theme_classic () +
    scale_color_manual(values=c("Positive"="palegreen1","Negative"="red1","No significant"="grey50"),
                   labels = c("Positive","Negative","No significant"))+
    labs(title="", x="rho", y = "-log10(q.value)")+
    geom_text(x=0, y=ymax-5, label="CX3", size=3, colour="black")+
    ylim(0,ymax)+
    xlim(xmin, xmax)+
    coord_capped_cart(left = "both", bottom = "both") +
    theme(axis.title.y = element_text(margin = margin(r=0.1, unit = "cm"),size=6))+
    theme(axis.title.x = element_text(margin = margin(t=0.1, unit = "cm"),size=6))+
    theme(axis.text.x = element_text(colour = "black", size = 6))+
    theme(axis.text.y = element_text(colour = "black", size = 6))+
    theme(title = element_text(face="bold", colour = "black", size = 7))+
    theme(plot.margin = margin(t = 0.2, r = 0.2, b = 0.2, l = 0.2, unit = "cm"))
  
pp1 <- ggarrange(p1, p2, p3, align = "h", ncol = 3, nrow = 1)

#HR 
path   <- "HR"
genes  <- DNARepairMechanisms$gene[DNARepairMechanisms$DNARepairMechanism %in% path]
cor    <- correlations[correlations$gene %in% genes,]
cor    <- mutate(cor, significant=ifelse(abs(cor$rho)<0.1, "No significant", 
                                           ifelse(abs(cor$rho)>0.1 & cor$rho>0, "Positive",
                                                  "Negative")))
p1 <- ggplot(cor[cor$signature=="CX2",], aes(x=as.numeric(rho), y=-log10(q.value), label = gene)) +      
    geom_point(aes(fill=significant, color=significant), show.legend=F, size=1)+
    geom_text_repel(aes(label = ifelse(-log10(q.value) > 30, gene, "")), size=1.8, max.overlaps = Inf,min.segment.length = 0)+
    theme_classic () +
    scale_color_manual(values=c("Positive"="palegreen1","Negative"="red1","No significant"="grey50"),
                   labels = c("Positive","Negative","No significant"))+
    labs(title=path, x="rho", y = "-log10(q.value)")+
    geom_text(x=0, y=ymax-5, label="CX2", size=3, colour="black")+
    ylim(0,ymax)+
    xlim(xmin, xmax)+
    coord_capped_cart(left = "both", bottom = "both") +
    theme(axis.title.y = element_text(margin = margin(r=0.1, unit = "cm"),size=6))+
    theme(axis.title.x = element_text(margin = margin(t=0.1, unit = "cm"),size=6))+
    theme(axis.text.x = element_text(colour = "black", size = 6))+
    theme(axis.text.y = element_text(colour = "black", size = 6))+
    theme(title = element_text(face="bold", colour = "black", size = 7))+
    theme(plot.margin = margin(t = 0.2, r = 0.2, b = 0.2, l = 0.2, unit = "cm"))

p2 <- ggplot(cor[cor$signature=="CX5",], aes(x=as.numeric(rho), y=-log10(q.value), label = gene)) +      
    geom_point(aes(fill=significant, color=significant), show.legend=F, size=1)+
    geom_text_repel(aes(label = ifelse(-log10(q.value) > 30, gene, "")), size=1.8, max.overlaps = Inf,min.segment.length = 0)+
    theme_classic () +
    scale_color_manual(values=c("Positive"="palegreen1","Negative"="red1","No significant"="grey50"),
                   labels = c("Positive","Negative","No significant"))+
    labs(title="", x="rho", y = "-log10(q.value)")+
    geom_text(x=0, y=ymax-5, label="CX5", size=3, colour="black")+
    ylim(0,ymax)+
    xlim(xmin, xmax)+
    coord_capped_cart(left = "both", bottom = "both") +
    theme(axis.title.y = element_text(margin = margin(r=0.1, unit = "cm"),size=6))+
    theme(axis.title.x = element_text(margin = margin(t=0.1, unit = "cm"),size=6))+
    theme(axis.text.x = element_text(colour = "black", size = 6))+
    theme(axis.text.y = element_text(colour = "black", size = 6))+
    theme(title = element_text(face="bold", colour = "black", size = 7))+
    theme(plot.margin = margin(t = 0.2, r = 0.2, b = 0.2, l = 0.2, unit = "cm"))

p3 <- ggplot(cor[cor$signature=="CX3",], aes(x=as.numeric(rho), y=-log10(q.value), label = gene)) +
    geom_point(aes(fill=significant, color=significant), show.legend=F, size=1)+
    geom_text_repel(aes(label = ifelse(-log10(q.value) > 35, gene, "")), size=1.8, max.overlaps = Inf, min.segment.length = 0)+
    theme_classic () +
    scale_color_manual(values=c("Positive"="palegreen1","Negative"="red1","No significant"="grey50"),
                   labels = c("Positive","Negative","No significant"))+
    labs(title="", x="rho", y = "-log10(q.value)")+
    geom_text(x=0, y=ymax-5, label="CX3", size=3, colour="black")+
    ylim(0,ymax)+
    xlim(xmin, xmax)+
    coord_capped_cart(left = "both", bottom = "both") +
    theme(axis.title.y = element_text(margin = margin(r=0.1, unit = "cm"),size=6))+
    theme(axis.title.x = element_text(margin = margin(t=0.1, unit = "cm"),size=6))+
    theme(axis.text.x = element_text(colour = "black", size = 6))+
    theme(axis.text.y = element_text(colour = "black", size = 6))+
    theme(title = element_text(face="bold", colour = "black", size = 7))+
    theme(plot.margin = margin(t = 0.2, r = 0.2, b = 0.2, l = 0.2, unit = "cm"))
  
pp2 <- ggarrange(p1, p2, p3, align = "h", ncol = 3, nrow = 1)

#MMR 
path   <- "MMR"
genes  <- DNARepairMechanisms$gene[DNARepairMechanisms$DNARepairMechanism %in% path]
cor    <- correlations[correlations$gene %in% genes,]
cor    <- mutate(cor, significant=ifelse(abs(cor$rho)<0.1, "No significant", 
                                           ifelse(abs(cor$rho)>0.1 & cor$rho>0, "Positive",
                                                  "Negative")))
p1 <- ggplot(cor[cor$signature=="CX2",], aes(x=as.numeric(rho), y=-log10(q.value), label = gene)) +      
    geom_point(aes(fill=significant, color=significant), show.legend=F, size=1)+
    geom_text_repel(aes(label = ifelse(-log10(q.value) > 25, gene, "")), size=1.8, max.overlaps = Inf,min.segment.length = 0)+
    theme_classic () +
    scale_color_manual(values=c("Positive"="palegreen1","Negative"="red1","No significant"="grey50"),
                   labels = c("Positive","Negative","No significant"))+
    labs(title=path, x="rho", y = "-log10(q.value)")+
    geom_text(x=0, y=ymax-5, label="CX2", size=3, colour="black")+
    ylim(0,ymax)+
    xlim(xmin, xmax)+
    coord_capped_cart(left = "both", bottom = "both") +
    theme(axis.title.y = element_text(margin = margin(r=0.1, unit = "cm"),size=6))+
    theme(axis.title.x = element_text(margin = margin(t=0.1, unit = "cm"),size=6))+
    theme(axis.text.x = element_text(colour = "black", size = 6))+
    theme(axis.text.y = element_text(colour = "black", size = 6))+
    theme(title = element_text(face="bold", colour = "black", size = 7))+
    theme(plot.margin = margin(t = 0.2, r = 0.2, b = 0.2, l = 0.2, unit = "cm"))

p2 <- ggplot(cor[cor$signature=="CX5",], aes(x=as.numeric(rho), y=-log10(q.value), label = gene)) +      
    geom_point(aes(fill=significant, color=significant), show.legend=F, size=1)+
    geom_text_repel(aes(label = ifelse(-log10(q.value) > 25, gene, "")), size=1.8, max.overlaps = Inf,min.segment.length = 0)+
    theme_classic () +
    scale_color_manual(values=c("Positive"="palegreen1","Negative"="red1","No significant"="grey50"),
                   labels = c("Positive","Negative","No significant"))+
    labs(title="", x="rho", y = "-log10(q.value)")+
    geom_text(x=0, y=ymax-5, label="CX5", size=3, colour="black")+
    ylim(0,ymax)+
    xlim(xmin, xmax)+
    coord_capped_cart(left = "both", bottom = "both") +
    theme(axis.title.y = element_text(margin = margin(r=0.1, unit = "cm"),size=6))+
    theme(axis.title.x = element_text(margin = margin(t=0.1, unit = "cm"),size=6))+
    theme(axis.text.x = element_text(colour = "black", size = 6))+
    theme(axis.text.y = element_text(colour = "black", size = 6))+
    theme(title = element_text(face="bold", colour = "black", size = 7))+
    theme(plot.margin = margin(t = 0.2, r = 0.2, b = 0.2, l = 0.2, unit = "cm"))

p3 <- ggplot(cor[cor$signature=="CX3",], aes(x=as.numeric(rho), y=-log10(q.value), label = gene)) +
    geom_point(aes(fill=significant, color=significant), show.legend=F, size=1)+
    geom_text_repel(aes(label = ifelse(-log10(q.value) > 25, gene, "")), size=1.8, max.overlaps = Inf,min.segment.length = 0)+
    theme_classic () +
    scale_color_manual(values=c("Positive"="palegreen1","Negative"="red1","No significant"="grey50"),
                   labels = c("Positive","Negative","No significant"))+
    labs(title="", x="rho", y = "-log10(q.value)")+
    geom_text(x=0, y=ymax-5, label="CX3", size=3, colour="black")+
    ylim(0,ymax)+
    xlim(xmin, xmax)+
    coord_capped_cart(left = "both", bottom = "both") +
    theme(axis.title.y = element_text(margin = margin(r=0.1, unit = "cm"),size=6))+
    theme(axis.title.x = element_text(margin = margin(t=0.1, unit = "cm"),size=6))+
    theme(axis.text.x = element_text(colour = "black", size = 6))+
    theme(axis.text.y = element_text(colour = "black", size = 6))+
    theme(title = element_text(face="bold", colour = "black", size = 7))+
    theme(plot.margin = margin(t = 0.2, r = 0.2, b = 0.2, l = 0.2, unit = "cm"))
  
pp3 <- ggarrange(p1, p2, p3, align = "h", ncol = 3, nrow = 1)

#NHEJ 
path   <- "NHEJ"
genes  <- DNARepairMechanisms$gene[DNARepairMechanisms$DNARepairMechanism %in% path]
cor    <- correlations[correlations$gene %in% genes,]
cor    <- mutate(cor, significant=ifelse(abs(cor$rho)<0.1, "No significant", 
                                           ifelse(abs(cor$rho)>0.1 & cor$rho>0, "Positive",
                                                  "Negative")))
p1 <- ggplot(cor[cor$signature=="CX2",], aes(x=as.numeric(rho), y=-log10(q.value), label = gene)) +      
    geom_point(aes(fill=significant, color=significant), show.legend=F, size=1)+
    geom_text_repel(aes(label = ifelse(-log10(q.value) > 25, gene, "")), size=1.8, max.overlaps = Inf,min.segment.length = 0)+
    theme_classic () +
    scale_color_manual(values=c("Positive"="palegreen1","Negative"="red1","No significant"="grey50"),
                   labels = c("Positive","Negative","No significant"))+
    labs(title=path, x="rho", y = "-log10(q.value)")+
    geom_text(x=0, y=ymax-5, label="CX2", size=3, colour="black")+
    ylim(0,ymax)+
    xlim(xmin, xmax)+
    coord_capped_cart(left = "both", bottom = "both") +
    theme(axis.title.y = element_text(margin = margin(r=0.1, unit = "cm"),size=6))+
    theme(axis.title.x = element_text(margin = margin(t=0.1, unit = "cm"),size=6))+
    theme(axis.text.x = element_text(colour = "black", size = 6))+
    theme(axis.text.y = element_text(colour = "black", size = 6))+
    theme(title = element_text(face="bold", colour = "black", size = 7))+
    theme(plot.margin = margin(t = 0.2, r = 0.2, b = 0.2, l = 0.2, unit = "cm"))

p2 <- ggplot(cor[cor$signature=="CX5",], aes(x=as.numeric(rho), y=-log10(q.value), label = gene)) +      
    geom_point(aes(fill=significant, color=significant), show.legend=F, size=1)+
    geom_text_repel(aes(label = ifelse(-log10(q.value) > 25, gene, "")), size=1.8, max.overlaps = Inf,min.segment.length = 0)+
    theme_classic () +
    scale_color_manual(values=c("Positive"="palegreen1","Negative"="red1","No significant"="grey50"),
                   labels = c("Positive","Negative","No significant"))+
    labs(title="", x="rho", y = "-log10(q.value)")+
    geom_text(x=0, y=ymax-5, label="CX5", size=3, colour="black")+
    ylim(0,ymax)+
    xlim(xmin, xmax)+
    coord_capped_cart(left = "both", bottom = "both") +
    theme(axis.title.y = element_text(margin = margin(r=0.1, unit = "cm"),size=6))+
    theme(axis.title.x = element_text(margin = margin(t=0.1, unit = "cm"),size=6))+
    theme(axis.text.x = element_text(colour = "black", size = 6))+
    theme(axis.text.y = element_text(colour = "black", size = 6))+
    theme(title = element_text(face="bold", colour = "black", size = 7))+
    theme(plot.margin = margin(t = 0.2, r = 0.2, b = 0.2, l = 0.2, unit = "cm"))

p3 <- ggplot(cor[cor$signature=="CX3",], aes(x=as.numeric(rho), y=-log10(q.value), label = gene)) +
    geom_point(aes(fill=significant, color=significant), show.legend=F, size=1)+
    geom_text_repel(aes(label = ifelse(-log10(q.value) > 25, gene, "")), size=1.8, max.overlaps = Inf,min.segment.length = 0)+
    theme_classic () +
    scale_color_manual(values=c("Positive"="palegreen1","Negative"="red1","No significant"="grey50"),
                   labels = c("Positive","Negative","No significant"))+
    labs(title="", x="rho", y = "-log10(q.value)")+
    geom_text(x=0, y=ymax-5, label="CX3", size=3, colour="black")+
    ylim(0,ymax)+
    xlim(xmin, xmax)+
    coord_capped_cart(left = "both", bottom = "both") +
    theme(axis.title.y = element_text(margin = margin(r=0.1, unit = "cm"),size=6))+
    theme(axis.title.x = element_text(margin = margin(t=0.1, unit = "cm"),size=6))+
    theme(axis.text.x = element_text(colour = "black", size = 6))+
    theme(axis.text.y = element_text(colour = "black", size = 6))+
    theme(title = element_text(face="bold", colour = "black", size = 7))+
    theme(plot.margin = margin(t = 0.2, r = 0.2, b = 0.2, l = 0.2, unit = "cm"))
  
pp4 <- ggarrange(p1, p2, p3, align = "h", ncol = 3, nrow = 1)

#SSA
path   <- "SSA"
genes  <- DNARepairMechanisms$gene[DNARepairMechanisms$DNARepairMechanism %in% path]
cor    <- correlations[correlations$gene %in% genes,]
cor    <- mutate(cor, significant=ifelse(abs(cor$rho)<0.1, "No significant", 
                                           ifelse(abs(cor$rho)>0.1 & cor$rho>0, "Positive",
                                                  "Negative")))
p1 <- ggplot(cor[cor$signature=="CX2",], aes(x=as.numeric(rho), y=-log10(q.value), label = gene)) +      
    geom_point(aes(fill=significant, color=significant), show.legend=F, size=1)+
    geom_text_repel(aes(label = ifelse(-log10(q.value) > 30, gene, "")), size=1.8, max.overlaps = Inf,min.segment.length = 0)+
    theme_classic () +
    scale_color_manual(values=c("Positive"="palegreen1","Negative"="red1","No significant"="grey50"),
                   labels = c("Positive","Negative","No significant"))+
    labs(title=path, x="rho", y = "-log10(q.value)")+
    geom_text(x=0, y=ymax-5, label="CX2", size=3, colour="black")+
    ylim(0,ymax)+
    xlim(xmin, xmax)+
    coord_capped_cart(left = "both", bottom = "both") +
    theme(axis.title.y = element_text(margin = margin(r=0.1, unit = "cm"),size=6))+
    theme(axis.title.x = element_text(margin = margin(t=0.1, unit = "cm"),size=6))+
    theme(axis.text.x = element_text(colour = "black", size = 6))+
    theme(axis.text.y = element_text(colour = "black", size = 6))+
    theme(title = element_text(face="bold", colour = "black", size = 7))+
    theme(plot.margin = margin(t = 0.2, r = 0.2, b = 0.2, l = 0.2, unit = "cm"))

p2 <- ggplot(cor[cor$signature=="CX5",], aes(x=as.numeric(rho), y=-log10(q.value), label = gene)) +      
    geom_point(aes(fill=significant, color=significant), show.legend=F, size=1)+
    geom_text_repel(aes(label = ifelse(-log10(q.value) > 30, gene, "")), size=1.8, max.overlaps = Inf,min.segment.length = 0)+
    theme_classic () +
    scale_color_manual(values=c("Positive"="palegreen1","Negative"="red1","No significant"="grey50"),
                   labels = c("Positive","Negative","No significant"))+
    labs(title="", x="rho", y = "-log10(q.value)")+
    geom_text(x=0, y=ymax-5, label="CX5", size=3, colour="black")+
    ylim(0,ymax)+
    xlim(xmin, xmax)+
    coord_capped_cart(left = "both", bottom = "both") +
    theme(axis.title.y = element_text(margin = margin(r=0.1, unit = "cm"),size=6))+
    theme(axis.title.x = element_text(margin = margin(t=0.1, unit = "cm"),size=6))+
    theme(axis.text.x = element_text(colour = "black", size = 6))+
    theme(axis.text.y = element_text(colour = "black", size = 6))+
    theme(title = element_text(face="bold", colour = "black", size = 7))+
    theme(plot.margin = margin(t = 0.2, r = 0.2, b = 0.2, l = 0.2, unit = "cm"))

p3 <- ggplot(cor[cor$signature=="CX3",], aes(x=as.numeric(rho), y=-log10(q.value), label = gene)) +
    geom_point(aes(fill=significant, color=significant), show.legend=F, size=1)+
    geom_text_repel(aes(label = ifelse(-log10(q.value) > 30, gene, "")), size=1.8, max.overlaps = Inf,min.segment.length = 0)+
    theme_classic () +
    scale_color_manual(values=c("Positive"="palegreen1","Negative"="red1","No significant"="grey50"),
                   labels = c("Positive","Negative","No significant"))+
    labs(title="", x="rho", y = "-log10(q.value)")+
    geom_text(x=0, y=ymax-5, label="CX3", size=3, colour="black")+
    ylim(0,ymax)+
    xlim(xmin, xmax)+
    coord_capped_cart(left = "both", bottom = "both") +
    theme(axis.title.y = element_text(margin = margin(r=0.1, unit = "cm"),size=6))+
    theme(axis.title.x = element_text(margin = margin(t=0.1, unit = "cm"),size=6))+
    theme(axis.text.x = element_text(colour = "black", size = 6))+
    theme(axis.text.y = element_text(colour = "black", size = 6))+
    theme(title = element_text(face="bold", colour = "black", size = 7))+
    theme(plot.margin = margin(t = 0.2, r = 0.2, b = 0.2, l = 0.2, unit = "cm"))
 
pp5 <- ggarrange(p1, p2, p3, align = "h", ncol = 3, nrow = 1)


#Volcano plot
pall <- ggarrange(pp, pp1, pp2, pp3, pp4, pp5, align = "v", ncol = 1, nrow = 6)

#sgv
ggsave(paste0(OUT,"/SuppFig_28_VolcanoPlot_nonNERPathways.svg"), pall, width = 180, height = 210, units = "mm")

# #pdf
# pdf(paste0(OUT,"/SupplFig_28_VolcanoPlots_non-NERPathways.pdf"), width = 180/25.4, height = 260/25.4)
# pall
# dev.off()

```
